---
title: "flexi_filter"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{flexi_filter}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(iidda.analysis)
```

# What is `flexi_filter` for?
The International Infectious Disease Data Archive (IIDDA) api is designed to
allow users to obtain filtered subsets of the data. The function
`iidda.analysis::flexi_filter` enables filtering IIDDA datasets with regex
queries. `flexi_filter` is [currently] bare-bones; searches that involve
filtering on more than one field simultaneously usually have to be be handled by
joining the results of two or more separate calls to `flexi_filter`.

# What does `flexi_filter` return?



# What you can search

`flexi_filter` can query any of the fields from different types of IIDDA
datasets. The type of dataset queried is specified with the argument
`response_type`, and can take the values `"metadata"` for metadata headers for
each flat datafile, `"columns"` for field names in the flat data files
themselves, or `"data_dictionary"` for definitions of the field names in either
the flat data files or their associated metadata. 

  


